---
title: "p8105_hw2_xc2474"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```


# Problem 1
## Read data

import csv file

```{r}
nyc_sub = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

clean the data

```{r}
nyc_sub_tidy = nyc_sub %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

This dataset contains `r ncol(nyc_sub_tidy)` variables, which is `r names(nyc_sub_tidy)`

My data cleaning steps are as follows:
1. import the data.
2. apply janitor::clean_names() function to change the variable names into a more readable way (all lowercase and only contain "_").
3. retain the variables we want using select funtion, which will remove the variables we don't need.
4. use recode() function to change all "YES" and "NO" in entry variable into correspongding "TRUE" and "FALSE" (make entry variable to a logical variable), and use mutate() function to retain the change in nyc_sub_tidy dataset.

The dimension of the resulting dataset is `r nrow(nyc_sub_tidy)` rows x `r ncol(nyc_sub_tidy)`

I think although these data are tidier than before, but they are not tidy enough.

## Question1

Q1: How many distinct stations are there?

```{r}
distinct(nyc_sub_tidy, station_name, line) %>%
  nrow()
```

A: There are 465 distinct stations.

Q2: How many stations are ADA compliant?

Here we define "ADA compliant" in a broad way, which means as long as ada=TRUE, we consider this station as ADA compliant:
```{r}
nyc_sub_tidy %>%
  filter(ada == TRUE) %>%
  nrow()
```
There are 468 stations are ADA compliant.

Q3: What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(filter(nyc_sub_tidy, vending == "NO")) / nrow(nyc_sub_tidy)
```

A: The proportion of station entrances / exits without vending allow entrance is 0.09796574.

Reformat data so that route number and route name are distinct variables

```{r}
nyc_sub_tidy = nyc_sub_tidy %>%
  gather(key = "route_number", value = "route_name", route1:route11) %>%
  separate(route_number, into = c('remove', 'route_number'), sep = 5) %>%
  select(-remove)
```

Q: How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
nyc_sub_tidy %>%
  filter(route_name == "A") %>%
  distinct(station_name, line) %>%
  nrow()
```

```{r}
nyc_sub_tidy %>%
  filter(route_name == "A", ada == TRUE) %>%
  distinct(station_name, line) %>%
  nrow()
```

A: There are 60 distinct station serve the A train, and there are 17 of them are ADA compliant.


# Problem 2

import Mr. Trash Wheel xlsx file, and clean the data:

1) specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
2) use reasonable variable names
3) omit rows that do not include dumpster-specific data
4)rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)


```{r}
tw_data = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = 'A2:N258' ) %>%
  janitor::clean_names() %>%
  filter(dumpster != "NA") %>%
  mutate(sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year.

```{r}
data_pre_2016 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = '2016 Precipitation', range = "A2:B14") %>%
  janitor::clean_names() %>%
  filter(total != 'NA') %>%
  mutate(year = 2016)

data_pre_2017 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = '2017 Precipitation', range = "A2:B14") %>%
  janitor::clean_names() %>%
  filter(total != 'NA') %>%
  mutate(year = 2017)
```

Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
 
```{r}
data_pre = bind_rows(data_pre_2016, data_pre_2017) %>%
  mutate(month = month.name[month])
```

In the Mr. Trash Wheel dataset, there are `r ncol(tw_data)` variables and and `r nrow(tw_data)` observations. According to the introduction of Mr. Trash Wheel, the key variables here are `dumpster`, `date`, `weight_tons` and `homes_powered`. The median number of sports balls in a dumpster in 2016 is `r median(filter(tw_data, year == 2016)$sports_balls)`. In the precipitation dataset, there are `r ncol(data_pre)` variables and `r nrow(data_pre)`. From these 20 observations, `r nrow(data_pre_2016)` observations come from 2016, `r nrow(data_pre_2017)` observations come from 2017. The key variables here are `month`, `year` and `total`. The total precipitation in 2017 was `r sum(data_pre_2017$total)`.

 
