---
title: "p8105_hw2_xc2474"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```


# Problem 1
## Read data

import csv file

```{r}
nyc_sub = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

clean the data

```{r}
nyc_sub_tidy = nyc_sub %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada, ada_notes) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

This dataset contains `r ncol(nyc_sub_tidy)` variables, which is `r names(nyc_sub_tidy)`

My data cleaning steps are as follows:
1. import the data.
2. apply janitor::clean_names() function to change the variable names into a more readable way (all lowercase and only contain "_").
3. retain the variables we want using select funtion, which will remove the variables we don't need.
4. use recode() function to change all "YES" and "NO" in entry variable into correspongding "TRUE" and "FALSE" (make entry variable to a logical variable), and use mutate() function to retain the change in nyc_sub_tidy dataset.

The dimension of the resulting dataset is `r nrow(nyc_sub_tidy)` rows x `r ncol(nyc_sub_tidy)`

I think although these data are tidier than before, but they are not tidy enough.

## Question1

Q1: How many distinct stations are there?

```{r}
distinct(nyc_sub_tidy, station_name, line) %>%
  nrow()
```

A: There are 465 distinct stations.

Q2: How many stations are ADA compliant?

1) If we define "ADA compliant" in a broad way, which means as long as ada=TRUE, we consider this station as ADA compliant:
```{r}
nyc_sub_tidy %>%
  filter(ada == TRUE) %>%
  nrow()
```
There are 468 stations are ADA compliant.


2) However, if we define "ADA compliant" in a way that only the ADA service is available is "ADA compliant":

```{r}
nyc_sub_tidy %>%
  filter(ada == TRUE, is.na(ada_notes)) %>%
  nrow()
```
??????WTF

Q3: What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(filter(nyc_sub_tidy, vending == "NO")) / nrow(nyc_sub_tidy)
```

A: The proportion of station entrances / exits without vending allow entrance is 0.09796574.




