---
title: "p8105_hw2_xc2474"
output: github_document
author: "Xinlei Chen"
date: 2018-10-01
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```


# Problem 1
## Read and clean data

import and clean the data

```{r}
nyc_sub_tidy = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

This dataset contains `r ncol(nyc_sub_tidy)` variables, which is `r names(nyc_sub_tidy)`

My data cleaning steps are as follows:

1. import the data.

2. apply janitor::clean_names() function to change the variable names into a more readable way (all lowercase and only contain "_").

3. retain the variables we want using select funtion, which will remove the variables we don't need.

4. use recode() function to change all "YES" and "NO" in entry variable into correspongding "TRUE" and "FALSE" (make entry variable to a logical variable), and use mutate() function to retain the change in nyc_sub_tidy dataset.


The dimension of the resulting dataset is `r nrow(nyc_sub_tidy)` rows x `r ncol(nyc_sub_tidy)` columns.

I think although these data are tidier than before, but they are not tidy enough, like columns from route1 to route11, it's not optimal.

## Questions

Q1: How many distinct stations are there?

```{r}
distinct(nyc_sub_tidy, station_name, line) %>%
  nrow()
```

A: There are 465 distinct stations.

Q2: How many stations are ADA compliant?

```{r}
nyc_sub_tidy %>%
  filter(ada == TRUE) %>%
  distinct(station_name, line) %>%
  nrow()
```
A: There are 84 stations are ADA compliant.

Q3: What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(filter(nyc_sub_tidy, vending == "NO", entry == "TRUE")) / 
  nrow(filter(nyc_sub_tidy, vending == "NO"))
```

A: The proportion of station entrances / exits without vending allow entrance is 0.3770492.

Reformat data so that route number and route name are distinct variables

```{r}
nyc_sub_tidy = nyc_sub_tidy %>%
  gather(key = "route_number", value = "route_name", route1:route11) %>%
  separate(route_number, into = c('remove', 'route_number'), sep = 5) %>%
  select(-remove)
```

Q4: How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
nyc_sub_tidy %>%
  filter(route_name == "A") %>%
  distinct(station_name, line) %>%
  nrow()
```

```{r}
nyc_sub_tidy %>%
  filter(route_name == "A", ada == TRUE) %>%
  distinct(station_name, line) %>%
  nrow()
```

A: There are 60 distinct station serve the A train, and there are 17 of them are ADA compliant.


# Problem 2
## Read and clean data

import Mr. Trash Wheel xlsx file, and clean the data:

1) specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
2) use reasonable variable names
3) omit rows that do not include dumpster-specific data
4)rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)


```{r}
tw_data = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", range = "A2:N338") %>%
  janitor::clean_names() %>%
  filter(dumpster != "NA") %>%
  mutate(sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year.

```{r}
data_pre_2016 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = '2016 Precipitation', range = "A2:B14") %>%
  janitor::clean_names() %>%
  filter(total != 'NA') %>%
  mutate(year = 2016)

data_pre_2017 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = '2017 Precipitation', range = "A2:B14") %>%
  janitor::clean_names() %>%
  filter(total != 'NA') %>%
  mutate(year = 2017)
```

Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
 
```{r}
data_pre = bind_rows(data_pre_2016, data_pre_2017) %>%
  mutate(month = month.name[month])
```

## Comment
In the Mr. Trash Wheel dataset, there are `r ncol(tw_data)` variables and and `r nrow(tw_data)` observations. According to the introduction of Mr. Trash Wheel, the key variables here are `dumpster`, `date`, `weight_tons` and `homes_powered`. The median number of sports balls in a dumpster in 2016 is `r median(filter(tw_data, year == 2016)$sports_balls)`. In the precipitation dataset, there are `r ncol(data_pre)` variables and `r nrow(data_pre)`. From these 20 observations, `r nrow(data_pre_2016)` observations come from 2016, `r nrow(data_pre_2017)` observations come from 2017. The key variables here are `month`, `year` and `total`. The total precipitation in 2017 was `r sum(data_pre_2017$total)`.

# Problem 3
## Read and clean data

```{r}
library(p8105.datasets)
```

Import and clean BRFSS data

1) format the data to use appropriate variable names;
2) focus on the “Overall Health” topic
3) exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
4) structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)
5) create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

```{r}
brfss_data = brfss_smart2010 %>%
  filter(Topic == 'Overall Health') %>%
  select(-c(Class, Topic, Question, Sample_Size, Confidence_limit_Low:GeoLocation)) %>%
  spread(key = Response, value = Data_value) %>%
  janitor::clean_names() %>%
  select(year, locationabbr, locationdesc, excellent, very_good, good, fair, poor) %>%
  mutate(high_prop = excellent + very_good)
```

## Questions

Q1: 

1) How many unique locations are included in the dataset?

```{r}
brfss_data %>%
  distinct(locationdesc) %>%
  nrow()
```
A: There are 404 unique locations are included in the dataset.

2) Is every state represented? 
```{r}
brfss_data %>%
  distinct(locationabbr) %>%
  nrow()
```
A: There are 51 unique states in this dataset, which equals to the number of states in the United States.

3)  What state is observed the most?
```{r}
count(brfss_data, locationabbr, sort = TRUE)
```
A: NJ

Q2: In 2002, what is the median of the “Excellent” response value?

```{r}
median(filter(brfss_data, year == 2002)$excellent, na.rm = TRUE)
```

A: Through removing NA value from the dataset, we get the median of "Excellent" response value which is 23.6. If we do not remove NA value from the dataset, the median will be NA.

Q3: Make a histogram of “Excellent” response values in the year 2002.

A: The histogram is shown as below:

```{r}
brfss_data %>%
  filter(year == 2002) %>%
  ggplot(aes(x = excellent)) + 
  geom_histogram()
```

Q4: Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.

A: As the range of year in the original dataset is exactly from 2002 to 2010, there is no need to filter the dataset according to the given period of the year. The scatterplot is shown as below:

```{r}
brfss_data %>%
  filter(locationdesc == 'NY - Queens County' | locationdesc == 'NY - New York County') %>%
  ggplot(aes(x = year, y = excellent)) +
  geom_point(aes(color = locationdesc))
```


(I'm really trying to commit more times in this assignment, so this time please please please don't deduct my points because of this)


