---
title: "p8105_hw2_xc2474"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```


#Problem 1

import csv file

```{r}
nyc_sub = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

clean the data

```{r}
nyc_sub_tidy = nyc_sub %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

This dataset contains 19 variables, which is line, station_name, stations_latitude, station longtitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, vending and ada. 

My data cleaning steps are as follows:
1. import the data.
2. apply janitor::clean_names() function to change the variable names into a more readable way (all lowercase and only contain "_").
3. retain the variables we want using select funtion, which will remove the variables we don't need.
4. use recode() function to change all "YES" and "NO" in entry variable into correspongding "TRUE" and "FALSE" (make entry variable to a logical variable), and use mutate() function to retain the change in nyc_sub_tidy dataset.

The dimension of the resulting dataset is `r nrow(nyc_sub_tidy)` rows x `r ncol(nyc_sub_tidy)`

I think although these data are tidier than before, but they are not tidy enough.









